Distributed Range-Only Localisation that Preserves Sensor and Navigator Privacies

### 2021-06-03 main.tex
Updated notation  - estimates are now with a hat  - all measurement variances are now time-dependent (also resulted in changing notation of the estimated modified measurement variance)  - made process model and process noise time-dependent as well (more general, no reason not to be)  - made the EKF prediction steps avoid k+1 notation (this led to confusion and some wrong equations that should be fixed now)  - Re-added series notation to pictures that was removed in the latest drat changes  - note that the large number of notation changes will need a thorough proof read! Picture colours updated

\subsection{Notation}
Throughout this work, we make use of the following notation. Underlined characters, $\vec{a}$, denote vectors, uppercase bold characters, $\mat{A}$, denote matrices. $\mat{A}^{-1}$ is the matrix inverse while $\mat{A}^\top$ is its transpose. The expected value of a random variable is denoted $\mathsf{E}[\cdot]$, while the variance of a random scalar and covariance of a random vector by $\mathsf{Var}[\cdot]$ and $\mathsf{Cov}[\cdot]$, respectively. $|a|$ denotes absolute value, $\lVert\vec{a}\rVert$ the vector norm, $\{\dots\}$ is used for sets and $\langle\dots\rangle$ for ordered sequences. When estimating a state, notation $\vec{x}_{k|l}$ will denote an estimate at timestep $k$ given measurements from timesteps up to and including $l$. When discussing cryptography, $\mathcal{E}_{k}(\cdot)$ and $\mathcal{D}_{k}(\cdot)$ will denote encryption and decryption with key $k$, respectively, with $k$ ommited when inferrable from context. $\mathbb{Z}_n$ is used for the set of integers m  odulo $n$ and $\mathbb{Z}^*_n$ for its multiplicative group. $\mathsf{lcm}(\cdot,\cdot)$ is the lowest common denominator, $\|$ the binary concatination operator and the term \textit{negligible function} refers to the cryptographic definition in \cite{katzIntroductionModernCryptography2008}.

@@

\section{Problem Statement} \label{sec:problem_statement}
In this work, we consider the context of privacy-preserving range-sensor navigation, where we want no sensor to learn information about the navigator or other sensors beyond their local measurements, and the navigator to learn no information about individual sensors beyond its location estimate. The problem is two-fold, in that we require explicit cryptographic requirements with a suitable encryption scheme meeting them, as well as an estimation scheme that can use the scheme in the context of range-sensor navigation.

@@

At every \textit{instance} $t$ (used to distinguish from an estimation \textit{timestep}), the navigator first broadcasts $m$ weights $\omega_j^{(t)}, 1 \leq j \leq m$ to all sensors $1\leq i\leq n$, who individually compute linear-combinations $l^{(t)}_i=\sum^m_{j=1}a_{j,i}^{(t)}\omega_i^{(t)}$ based on their measurement data $a_{j,i}$. Linear-combinations are then sent back to the navigator, who computes their sum $\sum^n_{i=1}l^{(t)}_{i}$. This two-step linear-combination aggregation protocol has been visually displayed in figure \ref{fig:agg_steps}.
\begin{figure}[htbp]
\centering
\begin{tikzpicture}[font=\footnotesize,scale=0.95]
    % Step 1
    \node at (3.25,5.5) {1. Broadcast and Combination Step};
    % Navigator
    \fill (3.25,4.5) [blue!40] ellipse (0.4 and 0.4);
    \pic[xscale=0.22,yscale=0.3] at (3.25,4.6725) {plane};
    % Sensors
    \node at (1.5,0.375) {Sensor $1$};
    \fill [red!40] (0.25,0.625) rectangle (2.875,1.625);
    \node at (1.625,1.125) {$\displaystyle l_1^{(t)} \!=\! \sum^m_{j=1}a_{1,j}^{(t)}\omega_j^{(t)}$};
    \node at (5,0.375) {Sensor $n$};
    \fill [red!40] (3.625,0.625) rectangle (6.25,1.625);
    \node at (5,1.125) {$\displaystyle l_n^{(t)} \!=\! \sum^m_{j=1}a_{n,j}^{(t)}\omega_j^{(t)}$};
    \fill [black] (3.5,1.125) circle (0.05);
    \fill [black] (3,1.125) circle (0.05);
    \fill [black] (3.25,1.125) circle (0.05);
    % Lines
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3.5,4) (5,1.875)};
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3,4) (1.5,1.875)};
    \fill [lightgray] (2.125,3.375) rectangle (4.375,2.625);
    \node at (3.25,3) {$\langle\omega_1^{(t)},\dots ,\omega_m^{(t)}\rangle$};

    % Step 2
    \node at (3.25,-0.75) {2. Aggregation Step};
    % Navigator
    \fill [blue!40] (1.375,-3) rectangle (5.125,-1.25);
    \pic[xscale=0.22,yscale=0.3] at (3.25,-1.4775) {plane};
    \node at (3.25,-2.375) {$\displaystyle \sum^{n}_{i=1}\sum^{m}_{j=1} a_{i,j}^{(t)}\omega_j^{(t)} = \sum^n_{i=1}l^{(t)}_{i}$};
    % Sensors
    \node at (1.25,-5.875) {Sensor $1$};
    \fill  (5.25,-5.25) [red!40] ellipse (0.4 and 0.4);
    \node at (5.25,-5.875) {Sensor $n$};
    \fill  (1.25,-5.25) [red!40] ellipse (0.4 and 0.4);
    \fill [black] (2.75,-5.25) circle (0.05);
    \fill [black] (3.75,-5.25) circle (0.05);
    \fill [black] (3.25,-5.25) circle (0.05);
    % Lines
    \draw [-latex] plot[smooth, tension=.7] coordinates {(5,-4.75) (3.75,-3.25)};
    \draw [-latex] plot[smooth, tension=.7] coordinates {(1.5,-4.75) (2.75,-3.25)};
    \fill [lightgray] (1.625,-3.75) rectangle (2.25,-4.25);
    \node at (2,-4) {$l_1^{(t)}$};
    \fill [lightgray] (4.25,-3.75) rectangle (4.875,-4.25);
    \node at (4.625,-4) {$l_n^{(t)}$};

    % Bounding rectangles
    \draw [gray] (0,6) rectangle (6.5,0);
    \draw [gray] (0,-0.25) rectangle (6.5,-6.25);
\end{tikzpicture}
\caption{Required linear-combination aggregation steps at instance $t$.}
\label{fig:agg_steps}
\end{figure}
In addition, we note that an alternative approach to the two-step protocol is computing $\sum^{m}_{j=1}(\omega_j^{(t)}\sum^{n}_{i=1} a_{i,j}^{(t)})$ at the navigator, requiring only values $a_{i,j}^{(t)}, 1 \leq j \leq m$ to be sent from each sensor $i$. We justify the use of bi-directional communication by reducing communication costs when the number of weights is larger than the number of sensors, $m>n$, and by sending fewer weights in the presence of repeats, as will be shown to be the case in section \ref{sec:priv_localization}.

@@

We are now ready to define the type of encryption scheme we want for the specified communication protocol and the security guarantees it should provide. We let a linear-combination aggregation scheme be defined as a tuple of the four algorithms $(\mathsf{Setup}, \mathsf{Enc}, \mathsf{CombEnc}, \mathsf{AggDec})$. These will be used by a trusted setup party, the navigator and sensors $i,\,1\leq i \leq n$. They are defined as follows.
\begin{LaTeXdescription}
    \item[$\mathsf{Setup}(\kappa)$] On input of security paramater $\kappa$, generate public parameters $\mathsf{pub}$, the number of weights $m$, the navigator's public and private keys $pk_0$ and $sk_0$ and the sensor private keys $sk_i,\,1\leq i \leq n$.
    \item[$\mathsf{Enc}(pk_0, x)$] The navigator and sensors can encrypt any value $x$ with the navigator's public key $pk_0$ and obtain the encryption $\mathcal{E}_{pk_0}(x)$.
    \item[$\mathsf{CombEnc}(t, pk_0, sk_i, \mathcal{E}_{pk_0}(\omega_1^{(t)}),\shrtdots,\mathcal{E}_{pk_0}(\omega_m^{(t)}), a^{(t)}_{i,1},\shrtdots,a^{(t)}_{i,m})$] At instance $t$, sensor $i$ computes and obtains the encrypted linear combination $l^{(t)}_i = \mathcal{E}_{pk_0,sk_i}(\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j)$ using its secret key $sk_i$.
    \item[$\mathsf{AggDec}(t, pk_0, sk_0, l^{(t)}_1,\shrtdots,l^{(t)}_n)$] At instance $t$, the navigator computes the aggregation of linear-combinations $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$ using its public and private keys $pk_0$, $sk_0$.
\end{LaTeXdescription}
The security notions we want the above algorithms to meet reflect the previously stated estimation privacy goals. The navigator should learn no information from individual sensors while sensors should learning no information from the navigator or any other sensors. In the context of the introduced communication protocol, this can be summarised as the following notions.
\begin{LaTeXdescription}
    \item[Indistinguishable Weights] No colluding subset of sensors gains any new knowledge about the navigator weights $\omega^{(t)}_j,\,1\leq j \leq m$ when receiving only their encryptions from the current and previous instances, and having the ability to encrypt plaintexts of their choice.
    \item[Linear-Combination Aggregator Obliviousness] No colluding subset \textit{excluding} the navigator gains additional information about the remaining sensor values to be weighted, $a^{(t)}_{i,j},\,1\leq j\leq m$, where sensor $i$ is not colluding, given only encryptions of their linear combinations $l_i$ from the current and previous instances. Any colluding subset \textit{including} the navigator learns only the sum of all linear combinations weighted by weights of their choice, $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$.
\end{LaTeXdescription}
While indistinguishable weights can be achieved by encrypting weights with an encryption scheme meeting the notion of Indistinguishability under the Chosen Ciphertext Attack (IND-CPA) \cite{katzIntroductionModernCryptography2008}, the novel notion of Linear-Combination Aggregator Obliviousness (LCAO) has been formalised as a typical cryptographic game between attacker and challenger in appendix \ref{app:lcao}. Lastly, we conclude the cryptographic problem definition with the following important remark.
\begin{remark}
    A leakage function including weights from the navigator requires extra care to be taken when giving its definition. If an attacker compromises the navigator, they have control over the weights, and therefore the leakage function. We note that in the leakage function above, $\sum^n_{i=1}\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j$, an individual sum weighted by the same weight may be learned by an attacker, \textit{e.g.}, $\sum^n_{i=1}a^{(t)}_{i,1}$ given weights $(1,0,\dots,0)$, but that individual sensor values $a^{(t)}_{i,j}$ remain private due to the assumption of a consistent broadcast.
\end{remark}

@@

\subsection{Estimation problem} \label{subsec:est_problem}
The estimation problem we consider, for which we will reduce communication to the protocol above, is localisation with range-only sensors. In this work, we will focus on the two-dimensional case for simplicity but will derive methods suitable for extension to a three-dimensional equivalent. The state that we wish to estimate captures the navigator position and the change in position ($\Delta$), and is given by
\begin{equation}
    \vec{x} = 
    \begin{bmatrix}
        x & \Delta x & y & \Delta y
    \end{bmatrix}^\top\,. \label{eqn:state_definition}
\end{equation}
The state evolves following a known system model, which at timestep $k$ is given by
\begin{equation}
    \vec{x}_k = \vec{f}(\vec{x}_{k-1})+\vec{w}_k\,, \label{eqn:system_model}
\end{equation}
with system noise covariance $\mathsf{Cov}[\vec{w}_k]=\mat{Q}_k$. The measurement model is dependent on sensor $i$'s location,
\begin{equation}
    \vec{s}_i = 
    \begin{bmatrix}
        s_{x,i} & s_{y,i}
    \end{bmatrix}^\top\,,
\end{equation} 
and is given by
\begin{equation}
    z_{k,i} = h_i(\vec{x}_k)+v_{k,i}\,, \label{eqn:measurement_model}
\end{equation}
with Gaussian measurement noises $v_{k,i} \sim \mathcal{N}(0,r_i)$ and measurement function $h_i$, for sensor $i$, as
\begin{equation}
    \begin{split}
        h_i(\vec{x}) &= \left\lVert
        \begin{bmatrix}
            x & y
        \end{bmatrix}^\top
        - \vec{s}_{i}\right\rVert \\
        &= \sqrt{(x-s_{x,i})^2 + (y-s_{y,i})^2}\,.
    \end{split}
\end{equation}

We aim to provide a filter that estimates the navigator's state $\vec{x}_k$, at every timestep $k$, without learning sensor positions $\vec{s}_i$, measurements $z_{k,i}$ and measurement variances $r_i$ beyond the information in the corresponding aggregation leakage function. Similarly, sensors should not learn any information about current state estimates or any other sensor information. Leakage will be further discussed in section \ref{subsec:leakage}, but we note that from any sequential state estimates, following known models, some sensor information leakage can be computed by the navigator. In the context of our leakage function, we will show that this corresponds to the global sums of private sensor information, while individual, or subsets of sensors', information remain private. Similarly, corrupted sensors with access to one or more measurements can produce state estimates of their own, leaking information about navigator state estimates, however, the most accurate estimates, requiring all measurements, will always remain private to the navigator.

@@

\section{Private Linear-Combination Aggregation Scheme} \label{sec:lcao_scheme}
In this section, we introduce an encryption scheme meeting the desired security properties in section \ref{subsec:crypto_problem}. The scheme is a combination of the Paillier and Joye-Libert schemes and provides encrypted weights meeting IND-CPA and encrypted aggregation meeting LCAO. Similarly to its constituents, the scheme bases its security on the DCRA and, as with the Joye-Libert scheme, requires a trusted party for initial key generation and distribution. The four algorithms defining our scheme are given as follows.
\begin{LaTeXdescription}
    \item[$\mathsf{Setup}(\kappa)$] On input parameter $\kappa$, generate two equal length, sufficiently large, primes $p$ and $q$, and compute $N=pq$. Define a hash function $H:\mathbb{Z} \rightarrow \mathbb{Z}_{N^2}^*$, choose a the number of weights to combine, $m>1$, and set public parameter $\mathsf{pub}=H$, navigator public key $pk_0 = N$ and navigator private key $sk_0=(p,q)$. Sensor secret keys are generated by choosing $sk_i,\,1\leq i\leq n-1$ uniformly from $\mathbb{Z}_{N^2}$ and setting the last key as $sk_n = -\sum^{n-1}_{i=1}sk_i \pmod{N^2}$.

@@

\section{Privacy-Preserving Localization} \label{sec:priv_localization}
With a concrete scheme meeting the LCAO notion, we can now put forward a localisation filter with communication that reduces to the required protocol. To produce an estimate of the state $\vec{x}_k$, we make use of an algebraic reformulation of the Extended Kalman Filter (EKF), the Extended Information Filter (EIF) \cite{maybeckStochasticModelsEstimation1982}, which reduces the filter update step to a single summation. The EIF update step requires the predicted state estimate $\vec{x}_{k|k-1}$ and estimate covariance $\mat{P}_{k|k-1}$ in the information vector and matrix forms
\begin{equation}
    \vec{y}_{k|k-1} = \mat{P}_{k|k-1}^{-1}\vec{x}_{k|k-1} \quad\textrm{and}\quad \mat{Y}_{k|k-1} = \mat{P}_{k|k-1}^{-1}\,,
\end{equation}
respectively. In this form, the update equations for $n$ sensor measurements at time $k$, given process and measurement models \eqref{eqn:system_model} and \eqref{eqn:measurement_model}, are given by
\begin{equation}
    \begin{split}
        &\vec{y}_{k|k} = \vec{y}_{k|k-1} + \\
        &\quad \sum^n_{i=1}\mat{H}^\top_{k,i} r^{-1}_i (z_{k,i} - h_i(\vec{x}_{k|k-1}) + \mat{H}_{k,i}\vec{x}_{k|k-1})
    \end{split} \label{eqn:eif_info_vec_update}
\end{equation}
and
\begin{equation}
    \mat{Y}_{k|k} = \mat{Y}_{k|k-1} + \sum^n_{i=1}\mat{H}^\top_{k,i} r^{-1}_i \mat{H}_{k,i}\,, \label{eqn:eif_info_mat_update}
\end{equation}
with Jacobians
\begin{equation}
    \mat{H}_{k,i} = \left.\frac{\partial h_i}{\partial \vec{x}}\right|_{\rvec{x}_{k|k-1}} \label{eqn:measurement_jacobian}
\end{equation}
for sensors $1 \leq i \leq n$. The filter's prediction step can be computed by converting information vector $\vec{y}_{k|k}$ and matrix $\mat{Y}_{k|k}$ back to state estimate and covariance 
\begin{equation}
    \vec{x}_{k|k} = \mat{Y}_{k|k}^{-1}\vec{y}_{k|k} \quad\textrm{and}\quad \mat{P}_{k|k} = \mat{Y}_{k|k}^{-1}\,,
\end{equation}
before using the normal EKF prediction equations
\begin{equation}
    \vec{x}_{k+1|k} = \vec{f}(\vec{x}_{k|k}) \quad\textrm{and}\quad \mat{P}_{k+1|k} = \mat{F}_k\mat{P}_{k|k}\mat{F}_k^\top + \mat{Q}_k\,, \label{eqn:ekf_prediction}
\end{equation}
with Jacobian
\begin{equation}
    \mat{F}_{k} = \left.\frac{\partial \vec{f}}{\partial \vec{x}}\right|_{\vec{x}_{k|k}}\,. \label{eqn:jacobian_f}
\end{equation}

>>>>

\subsection{Notation}
Throughout this work, we make use of the following notation. Underlined characters, $\vec{a}$, denote vectors, uppercase bold characters, $\mat{A}$, denote matrices. $\mat{A}^{-1}$ is the matrix inverse while $\mat{A}^\top$ is its transpose. The expected value of a random variable is denoted $\mathsf{E}[\cdot]$, while the variance of a random scalar and covariance of a random vector by $\mathsf{Var}[\cdot]$ and $\mathsf{Cov}[\cdot]$, respectively. $|a|$ denotes absolute value, $\lVert\vec{a}\rVert$ the vector norm, $\{\dots\}$ is used for sets and $\langle\dots\rangle$ for ordered sequences. When estimating a state, notation $\hat{\vec{x}}_{k|l}$ will denote an estimate of $\vec{x}$ at timestep $k$ given measurements from timesteps up to and including $l$. When discussing cryptography, $\mathcal{E}_{k}(\cdot)$ and $\mathcal{D}_{k}(\cdot)$ will denote encryption and decryption with key $k$, respectively, with $k$ ommited when inferrable from context. $\mathbb{Z}_n$ is used for the set of integers m  odulo $n$ and $\mathbb{Z}^*_n$ for its multiplicative group. $\mathsf{lcm}(\cdot,\cdot)$ is the lowest common denominator, $\|$ the binary concatination operator and the term \textit{negligible function} refers to the cryptographic definition in \cite{katzIntroductionModernCryptography2008}.

@@

\section{Problem Statement} \label{sec:problem_statement}
In this work, we consider the context of privacy-preserving range sensor navigation, where we want no sensor to learn information about the navigator or other sensors beyond their local measurements, and the navigator to learn no information about individual sensors beyond its location estimate. The problem is two-fold, in that we require explicit cryptographic requirements with a suitable encryption scheme meeting them, as well as an estimation scheme that can use the scheme in the context of range-only navigation.

@@

At every \textit{instance} $t$ (used to distinguish from an estimation \textit{timestep}), the navigator first broadcasts $m$ weights $\omega_j^{(t)}, 1 \leq j \leq m$ to all sensors $1\leq i\leq n$, who individually compute linear combinations $l^{(t)}_i=\sum^m_{j=1}a_{j,i}^{(t)}\omega_i^{(t)}$ based on their measurement data $a_{j,i}$. Linear combinations are then sent back to the navigator, who computes their sum $\sum^n_{i=1}l^{(t)}_{i}$. This two-step linear combination aggregation protocol has been visually displayed in figure \ref{fig:agg_steps}.
\begin{figure}[htbp]
\centering
\begin{tikzpicture}[font=\footnotesize,scale=0.95]
    % Step 1
    \node at (3.25,5.5) {1. Broadcast and Combination Step};
    % Navigator
    \fill (3.25,4.5) [green!80!black!60] ellipse (0.4 and 0.4);
    \pic[xscale=0.22,yscale=0.3] at (3.25,4.6725) {plane};
    % Sensors
    \node at (1.5,0.375) {Sensor $1$};
    \fill [blue!80!black!40] (0.25,0.625) rectangle (2.875,1.625);
    \node at (1.625,1.125) {$\displaystyle l_1^{(t)} \!=\! \sum^m_{j=1}a_{1,j}^{(t)}\omega_j^{(t)}$};
    \node at (5,0.375) {Sensor $n$};
    \fill [blue!80!black!40] (3.625,0.625) rectangle (6.25,1.625);
    \node at (5,1.125) {$\displaystyle l_n^{(t)} \!=\! \sum^m_{j=1}a_{n,j}^{(t)}\omega_j^{(t)}$};
    \fill [black] (3.5,1.125) circle (0.05);
    \fill [black] (3,1.125) circle (0.05);
    \fill [black] (3.25,1.125) circle (0.05);
    % Lines
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3.5,4) (5,1.875)};
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3,4) (1.5,1.875)};
    \fill [lightgray] (2.125,3.375) rectangle (4.375,2.625);
    \node at (3.25,3) {$\langle\omega_1^{(t)},\dots ,\omega_m^{(t)}\rangle$};

    % Step 2
    \node at (3.25,-0.75) {2. Aggregation Step};
    % Navigator
    \fill [green!80!black!60] (1.375,-3) rectangle (5.125,-1.25);
    \pic[xscale=0.22,yscale=0.3] at (3.25,-1.4775) {plane};
    \node at (3.25,-2.375) {$\displaystyle \sum^{n}_{i=1}\sum^{m}_{j=1} a_{i,j}^{(t)}\omega_j^{(t)} = \sum^n_{i=1}l^{(t)}_{i}$};
    % Sensors
    \node at (1.25,-5.875) {Sensor $1$};
    \fill  (5.25,-5.25) [blue!80!black!40] ellipse (0.4 and 0.4);
    \node at (5.25,-5.875) {Sensor $n$};
    \fill  (1.25,-5.25) [blue!80!black!40] ellipse (0.4 and 0.4);
    \fill [black] (2.75,-5.25) circle (0.05);
    \fill [black] (3.75,-5.25) circle (0.05);
    \fill [black] (3.25,-5.25) circle (0.05);
    % Lines
    \draw [-latex] plot[smooth, tension=.7] coordinates {(5,-4.75) (3.75,-3.25)};
    \draw [-latex] plot[smooth, tension=.7] coordinates {(1.5,-4.75) (2.75,-3.25)};
    \fill [lightgray] (1.625,-3.75) rectangle (2.25,-4.25);
    \node at (2,-4) {$l_1^{(t)}$};
    \fill [lightgray] (4.25,-3.75) rectangle (4.875,-4.25);
    \node at (4.625,-4) {$l_n^{(t)}$};

    % Bounding rectangles
    \draw [gray] (0,6) rectangle (6.5,0);
    \draw [gray] (0,-0.25) rectangle (6.5,-6.25);
\end{tikzpicture}
\caption{Required linear combination aggregation steps at instance $t$.}
\label{fig:agg_steps}
\end{figure}
In addition, we note that an alternative approach to the two-step protocol is computing $\sum^{m}_{j=1}(\omega_j^{(t)}\sum^{n}_{i=1} a_{i,j}^{(t)})$ at the navigator, requiring only values $a_{i,j}^{(t)}, 1 \leq j \leq m$ to be sent from each sensor $i$. We justify the use of bi-directional communication by reducing communication costs when the number of weights is larger than the number of sensors, $m>n$, and by sending fewer weights in the presence of repeats, as will be shown to be the case in section \ref{sec:priv_localization}.

@@

We are now ready to define the type of encryption scheme we want for the specified communication protocol and the security guarantees it should provide. We let a linear combination aggregation scheme be defined as a tuple of the four algorithms $(\mathsf{Setup}, \mathsf{Enc}, \mathsf{CombEnc}, \mathsf{AggDec})$. These will be used by a trusted setup party, the navigator and sensors $i,\,1\leq i \leq n$. They are defined as follows.
\begin{LaTeXdescription}
    \item[$\mathsf{Setup}(\kappa)$] On input of security paramater $\kappa$, generate public parameters $\mathsf{pub}$, the number of weights $m$, the navigator's public and private keys $pk_0$ and $sk_0$ and the sensor private keys $sk_i,\,1\leq i \leq n$.
    \item[$\mathsf{Enc}(pk_0, x)$] The navigator and sensors can encrypt any value $x$ with the navigator's public key $pk_0$ and obtain the encryption $\mathcal{E}_{pk_0}(x)$.
    \item[$\mathsf{CombEnc}(t, pk_0, sk_i, \mathcal{E}_{pk_0}(\omega_1^{(t)}),\shrtdots,\mathcal{E}_{pk_0}(\omega_m^{(t)}), a^{(t)}_{i,1},\shrtdots,a^{(t)}_{i,m})$] At instance $t$, sensor $i$ computes and obtains the encrypted linear combination $l^{(t)}_i = \mathcal{E}_{pk_0,sk_i}(\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j)$ using its secret key $sk_i$.
    \item[$\mathsf{AggDec}(t, pk_0, sk_0, l^{(t)}_1,\shrtdots,l^{(t)}_n)$] At instance $t$, the navigator computes the aggregation of linear combinations $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$ using its public and private keys $pk_0$, $sk_0$.
\end{LaTeXdescription}
The security notions we want the above algorithms to meet reflect the previously stated estimation privacy goals. The navigator should learn no information from individual sensors while sensors should learning no information from the navigator or any other sensors. In the context of the introduced communication protocol, this can be summarised as the following notions.
\begin{LaTeXdescription}
    \item[Indistinguishable Weights] No colluding subset of sensors gains any new knowledge about the navigator weights $\omega^{(t)}_j,\,1\leq j \leq m$ when receiving only their encryptions from the current and previous instances, and having the ability to encrypt plaintexts of their choice.
    \item[Linear Combination Aggregator Obliviousness] No colluding subset \textit{excluding} the navigator gains additional information about the remaining sensor values to be weighted, $a^{(t)}_{i,j},\,1\leq j\leq m$, where sensor $i$ is not colluding, given only encryptions of their linear combinations $l_i$ from the current and previous instances. Any colluding subset \textit{including} the navigator learns only the sum of all linear combinations weighted by weights of their choice, $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$.
\end{LaTeXdescription}
While indistinguishable weights can be achieved by encrypting weights with an encryption scheme meeting the notion of Indistinguishability under the Chosen Ciphertext Attack (IND-CPA) \cite{katzIntroductionModernCryptography2008}, the novel notion of Linear Combination Aggregator Obliviousness (LCAO) has been formalised as a typical cryptographic game between attacker and challenger in appendix \ref{app:lcao}. Lastly, we conclude the cryptographic problem definition with the following important remark.
\begin{remark}
    A leakage function including weights from the navigator requires extra care to be taken when giving its definition. If an attacker compromises the navigator, they have control over the weights, and therefore the leakage function. We note that in the leakage function above, $\sum^n_{i=1}\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j$, an individual sum weighted by the same weight may be learned by an attacker, \textit{e.g.}, $\sum^n_{i=1}a^{(t)}_{i,1}$ given weights $(1,0,\dots,0)$, but that individual sensor values $a^{(t)}_{i,j}$ remain private due to the assumption of a consistent broadcast.
\end{remark}

@@

\subsection{Estimation problem} \label{subsec:est_problem}
The estimation problem we consider, for which we will reduce communication to the protocol above, is localisation with range-only sensors. In this work, we will focus on the two-dimensional case for simplicity but will derive methods suitable for extension to a three-dimensional equivalent. The state that we wish to estimate captures the navigator position and the change in position ($\Delta$), and is given by
\begin{equation}
    \vec{x} = 
    \begin{bmatrix}
        x & \Delta x & y & \Delta y
    \end{bmatrix}^\top\,. \label{eqn:state_definition}
\end{equation}
The state evolves following a known system model, which at timestep $k$ is given by
\begin{equation}
    \vec{x}_k = \vec{f}_k(\vec{x}_{k-1})+\vec{w}_k\,, \label{eqn:system_model}
\end{equation}
with system noise covariance $\mathsf{Cov}[\vec{w}_k]=\mat{Q}_k$. Measurements of the system follow a measurement model dependent on sensor $i$, $0\leq i \leq n$, given by 
\begin{equation}
    z_{k,i} = h_i(\vec{x}_k)+v_{k,i}\,, \label{eqn:measurement_model}
\end{equation}
with Gaussian measurement noises $v_{k,i} \sim \mathcal{N}(0,r_{k,i})$ and measurement function
\begin{equation}
    \begin{split}
        h_i(\vec{x}) &= \left\lVert
        \begin{bmatrix}
            x & y
        \end{bmatrix}^\top
        - \vec{s}_{i}\right\rVert \\
        &= \sqrt{(x-s_{x,i})^2 + (y-s_{y,i})^2}\,,
    \end{split}
\end{equation}
where
\begin{equation}
    \vec{s}_i = 
    \begin{bmatrix}
        s_{x,i} & s_{y,i}
    \end{bmatrix}^\top
\end{equation} 
is the location of sensor $i$.

We aim to provide a filter that estimates the navigator's state $\vec{x}_k$, at every timestep $k$, without learning sensor positions $\vec{s}_i$, measurements $z_{k,i}$ and measurement variances $r_{k,i}$ beyond the information in the corresponding aggregation leakage function. Similarly, sensors should not learn any information about current state estimates or any other sensor information. Leakage will be further discussed in section \ref{subsec:leakage}, but we note that from any sequential state estimates, following known models, some sensor information leakage can be computed by the navigator. In the context of our leakage function, we will show that this corresponds to the global sums of private sensor information, while individual, or subsets of sensors', information remain private. Similarly, corrupted sensors with access to one or more measurements can produce state estimates of their own, leaking information about navigator state estimates, however, the most accurate estimates, requiring all measurements, will always remain private to the navigator.

@@

\section{Private Linear Combination Aggregation Scheme} \label{sec:lcao_scheme}
In this section, we introduce an encryption scheme meeting the desired security properties in section \ref{subsec:crypto_problem}. The scheme is a combination of the Paillier and Joye-Libert schemes and provides encrypted weights meeting IND-CPA and encrypted aggregation meeting LCAO. Similarly to its constituents, the scheme bases its security on the DCRA and, as with the Joye-Libert scheme, requires a trusted party for initial key generation and distribution. The four algorithms defining our scheme are given as follows.
\begin{LaTeXdescription}
    \item[$\mathsf{Setup}(\kappa)$] On input parameter $\kappa$, generate two equal length, sufficiently large, primes $p$ and $q$, and compute $N=pq$. Define a hash function $H:\mathbb{Z} \rightarrow \mathbb{Z}_{N^2}^*$, choose a the number of weights to combine, $m>1$, and set public parameter $\mathsf{pub}=H$, navigator public key $pk_0 = N$ and navigator private key $sk_0=(p,q)$. Sensor secret keys are generated by choosing $sk_i,\,1\leq i\leq n-1$ uniformly from $\mathbb{Z}_{N^2}$ and setting the last key as $sk_n = -\sum^{n-1}_{i=1}sk_i \pmod{N^2}$.

@@

\section{Privacy-Preserving Localization} \label{sec:priv_localization}
With a concrete scheme meeting the LCAO notion, we can now put forward a localisation filter with communication that reduces to the required protocol. To produce an estimate of the state $\vec{x}_k$, we make use of an algebraic reformulation of the Extended Kalman Filter (EKF), the Extended Information Filter (EIF) \cite{maybeckStochasticModelsEstimation1982}, which reduces the filter update step to a single summation. The EIF update step requires the predicted state estimate $\hat{\vec{x}}_{k|k-1}$ and estimate covariance $\mat{P}_{k|k-1}$ in the information vector and matrix forms
\begin{equation}
    \hat{\vec{y}}_{k|k-1} = \mat{P}_{k|k-1}^{-1}\hat{\vec{x}}_{k|k-1} \quad\textrm{and}\quad \mat{Y}_{k|k-1} = \mat{P}_{k|k-1}^{-1}\,,
\end{equation}
respectively. In this form, the update equations for $n$ sensor measurements at time $k$, given process and measurement models \eqref{eqn:system_model} and \eqref{eqn:measurement_model}, are given by
\begin{equation}
    \begin{split}
        &\hat{\vec{y}}_{k|k} = \hat{\vec{y}}_{k|k-1} + \\
        &\quad \sum^n_{i=1}\mat{H}^\top_{k,i} r^{-1}_i (z_{k,i} - h_i(\hat{\vec{x}}_{k|k-1}) + \mat{H}_{k,i}\hat{\vec{x}}_{k|k-1})
    \end{split} \label{eqn:eif_info_vec_update}
\end{equation}
and
\begin{equation}
    \mat{Y}_{k|k} = \mat{Y}_{k|k-1} + \sum^n_{i=1}\mat{H}^\top_{k,i} r^{-1}_i \mat{H}_{k,i}\,, \label{eqn:eif_info_mat_update}
\end{equation}
with Jacobians
\begin{equation}
    \mat{H}_{k,i} = \left.\frac{\partial h_i}{\partial \vec{x}}\right|_{\hat{\vec{x}}_{k|k-1}} \label{eqn:measurement_jacobian}
\end{equation}
for sensors $1 \leq i \leq n$. The filter's prediction step can be computed by converting information vector $\hat{\vec{y}}_{k|k}$ and matrix $\mat{Y}_{k|k}$ back to state estimate and covariance 
\begin{equation}
    \hat{\vec{x}}_{k|k} = \mat{Y}_{k|k}^{-1}\hat{\vec{y}}_{k|k} \quad\textrm{and}\quad \mat{P}_{k|k} = \mat{Y}_{k|k}^{-1}\,,
\end{equation}
before using the normal EKF prediction equations
\begin{equation}
    \hat{\vec{x}}_{k|k-1} \!=\! \vec{f}_k(\hat{\vec{x}}_{k-1|k-1}) \ \textrm{and}\ \mat{P}_{k|k-1} \!=\! \mat{F}_k\mat{P}_{k-1|k-1}\mat{F}_k^\top + \mat{Q}_k, \label{eqn:ekf_prediction}
\end{equation}
with Jacobian
\begin{equation}
    \mat{F}_{k} = \left.\frac{\partial \vec{f}_k}{\partial \vec{x}}\right|_{\hat{\vec{x}}_{k-1|k-1}}\,. \label{eqn:jacobian_f}
\end{equation}
